---
title: "preliminaryTasks"
output: html_document
---

# Preliminary Task 1: Augmenting HTML to include header
```{r, message = F}
require(tidyverse)
require(tidytext)
require(textstem)
require(rvest)
require(qdapRegex)
require(stopwords)
require(tokenizers)
library(caret)
library(tidymodels)
```

## Updating Pre-processing Functions

Update `parse_fn` to include headers

```{r}
# function to parse html and clean text
parse_fn <- function(.html){
  read_html(.html) %>%
    html_elements('p, h1, h2, h3, h4, h5, h6') %>%
    html_text2() %>%
    str_c(collapse = ' ') %>%
    rm_url() %>%
    rm_email() %>%
    str_remove_all('\'') %>%
    str_replace_all(paste(c('\n', 
                            '[[:punct:]]', 
                            'nbsp', 
                            '[[:digit:]]', 
                            '[[:symbol:]]'),
                          collapse = '|'), ' ') %>%
    str_replace_all("([a-z])([A-Z])", "\\1 \\2") %>%
    tolower() %>%
    str_replace_all("\\s+", " ")
}

# function to apply to claims data
parse_data <- function(.df){
  out <- .df %>%
    filter(str_detect(text_tmp, '<!')) %>%
    rowwise() %>%
    mutate(text_clean = parse_fn(text_tmp)) %>%
    unnest(text_clean) 
  return(out)
}

nlp_fn <- function(parse_data.out){
  out <- parse_data.out %>% 
    unnest_tokens(output = token, 
                  input = text_clean, 
                  token = 'words',
                  stopwords = str_remove_all(stop_words$word, 
                                             '[[:punct:]]')) %>%
    mutate(token.lem = lemmatize_words(token)) %>%
    filter(str_length(token.lem) > 2) %>%
    count(.id, bclass, token.lem, name = 'n') %>%
    bind_tf_idf(term = token.lem, 
                document = .id,
                n = n) %>%
    pivot_wider(id_cols = c('.id', 'bclass'),
                names_from = 'token.lem',
                values_from = 'tf_idf',
                values_fill = 0)
  return(out)
}

# clean_df <- claims_test %>%
#   parse_data() %>%
#   select(.id, bclass, text_clean)
#load('../data/claims-raw.RData')
#claims_clean <- claims_raw %>%
 # parse_data()

load('../data/claims-raw.RData')

# use a smaller subset for Task 1 and Task 2 to avoid crashing
set.seed(123)
claims_clean <- claims_raw %>%
  slice_sample(n = 400) %>%   # try 300–500; increase only if your machine handles it
  parse_data()
```

## Logistic PCR

# run not this chunk but the one below 
```{r}
text_features <- nlp_fn(claims_clean)
model_df <- text_features %>%
  select(-.id)

model_df$bclass <- as.factor(model_df$bclass)

set.seed(123)
split <- initial_split(model_df, strata = bclass)

train_df <- training(split)
test_df  <- testing(split)

pcr_recipe <- recipe(bclass ~ ., data = train_df) %>%
  step_zv(all_predictors()) %>% 
  step_normalize(all_predictors()) %>%
  step_pca(all_predictors(), num_comp = 50)

logistic_spec <- logistic_reg(
  mode = "classification",
  penalty = 0.01,
  mixture = 0) %>%
  set_engine("glmnet")

pcr_wf <- workflow() %>%
  add_recipe(pcr_recipe) %>%
  add_model(logistic_spec)

final_fit <- fit(pcr_wf, train_df)

# Predictions (class)
pred_class <- predict(final_fit, test_df)

# Predictions (probabilities)
pred_prob <- predict(final_fit, test_df, type = "prob")

# Combine with test data
bclass <- as_tibble(test_df$bclass)
results <- bind_cols(bclass, pred_class, pred_prob) %>% 
  rename(bclass = value)

# Accuracy
accuracy(results, truth = bclass, estimate = .pred_class)
# 0.616
```

# this one 
```{r}
text_features <- nlp_fn(claims_clean)
model_df <- text_features %>%
  select(-.id)
```

```{r}
model_df$bclass <- as.factor(model_df$bclass)

set.seed(123)
split <- initial_split(model_df, strata = bclass)

train_df <- training(split)
test_df  <- testing(split)

pcr_recipe <- recipe(bclass ~ ., data = train_df) %>%
  step_zv(all_predictors()) %>% 
  step_normalize(all_predictors()) %>%
  step_pca(all_predictors(), num_comp = 50)

logistic_spec <- logistic_reg(
  mode = "classification",
  penalty = 0.01,
  mixture = 0) %>%
  set_engine("glmnet")

pcr_wf <- workflow() %>%
  add_recipe(pcr_recipe) %>%
  add_model(logistic_spec)

set.seed(123)
folds <- vfold_cv(train_df, v = 5, strata = bclass)

pca_grid <- tibble(num_comp = c(10, 20, 50, 100, 200))

tune_res <- tune_grid(
  pcr_wf,
  resamples = folds,
  grid = pca_grid,
  metrics = metric_set(accuracy)
)

best_pcs <- select_best(tune_res, metric = "accuracy")
best_pcs

final_wf <- finalize_workflow(pcr_wf, best_pcs)

final_fit <- fit(final_wf, train_df)

# Predictions (class)
pred_class <- predict(final_fit, test_df)

# Predictions (probabilities)
pred_prob <- predict(final_fit, test_df, type = "prob")

# Combine with test data
bclass <- as_tibble(test_df$bclass)
results <- bind_cols(bclass, pred_class, pred_prob) %>% 
  rename(bclass = value)

# Accuracy
accuracy(results, truth = bclass, estimate = .pred_class)
#
```

## Preliminary task 2 

1. Bigram tokenization and TF–IDF

Add a bigram nlp function that parallels your nlp_fn():

```{r}
# --- Task 2: Bigram tokenization and TF–IDF ---

nlp_fn_bigrams <- function(parse_data.out) {
  parse_data.out %>%
    unnest_tokens(
      output = bigram,
      input  = text_clean,
      token  = "ngrams",
      n      = 2
    ) %>%
    # optional cleanup: drop bigrams with digits or punctuation
    filter(!str_detect(bigram, "[0-9]"),
           !str_detect(bigram, "[[:punct:]]")) %>%
    count(.id, bclass, bigram, name = "n") %>%
    bind_tf_idf(
      term     = bigram,
      document = .id,
      n        = n
    ) %>%
    pivot_wider(
      id_cols     = c(".id", "bclass"),
      names_from  = "bigram",
      values_from = "tf_idf",
      values_fill = 0
    )
}

# recompute word-level features to be explicit (safe even if already done)
text_features <- nlp_fn(claims_clean)

# bigram-level features
bigram_features <- nlp_fn_bigrams(claims_clean)


```

ˆ
2. Create a consistent train/test split by ID

2. Train/test split using text_features
```{r}
# IDs + label for splitting
set.seed(123)
split2 <- initial_split(text_features, strata = bclass)

train_text <- training(split2)
test_text  <- testing(split2)

# data frames for word-level modeling (drop .id)
train_word_df <- train_text %>% select(-.id)
test_word_df  <- test_text  %>% select(-.id)

train_word_df$bclass <- as.factor(train_word_df$bclass)
test_word_df$bclass  <- as.factor(test_word_df$bclass)


```
3. Word-only logistic PCR (baseline for Task 2)

```{r}
# --- Word-only logistic PCR (baseline) ---

# --- Word-only logistic PCR (baseline with tuning) ---

set.seed(123)
pcr_recipe_word <- recipe(bclass ~ ., data = train_word_df) %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_predictors()) %>%
  # make num_comp tunable instead of fixed at 50
  step_pca(all_predictors(), num_comp = tune())

logistic_spec <- logistic_reg(
  mode    = "classification",
  penalty = 0.01,
  mixture = 0
) %>%
  set_engine("glmnet")

pcr_wf_word <- workflow() %>%
  add_recipe(pcr_recipe_word) %>%
  add_model(logistic_spec)

set.seed(123)
folds_word <- vfold_cv(train_word_df, v = 5, strata = bclass)

# grid of PCA components to try
pca_grid <- tibble(num_comp = c(10, 20, 50, 100, 200))

tune_res_word <- tune_grid(
  pcr_wf_word,
  resamples = folds_word,
  grid      = pca_grid,
  metrics   = metric_set(accuracy)
)

best_pcs_word <- select_best(tune_res_word, metric = "accuracy")
best_pcs_word

final_wf_word <- finalize_workflow(pcr_wf_word, best_pcs_word)

final_fit_word <- fit(final_wf_word, train_word_df)

# predictions on test set (baseline)
pred_class_word <- predict(final_fit_word, test_word_df)

results_word <- bind_cols(
  bclass = test_word_df$bclass,
  pred_class_word
) %>%
  rename(bclass = bclass)

acc_word <- accuracy(results_word, truth = bclass, estimate = .pred_class)
acc_word


```

4. Get predicted log-odds (logit) from the word model

```{r}
# --- Predicted log-odds from word-only model ---

# probabilities for each class
train_prob_df <- predict(final_fit_word, train_word_df, type = "prob")
test_prob_df  <- predict(final_fit_word, test_word_df,  type = "prob")

# check column names if you're curious:
# names(train_prob_df)

# take the probability of the "positive" class = 2nd level of bclass
train_prob_word <- train_prob_df %>% dplyr::pull(2)
test_prob_word  <- test_prob_df  %>% dplyr::pull(2)

# safe logit function (avoid log(0))
logit <- function(p) {
  eps <- 1e-6
  p <- pmin(pmax(p, eps), 1 - eps)
  log(p / (1 - p))
}

train_logit_word <- logit(train_prob_word)
test_logit_word  <- logit(test_prob_word)
```

```{r}
# Sanity check that vector is not 0
length(train_logit_word)
nrow(train_text)
```


5. PCA on bigram TF–IDF and build stage-2 design matrices

```{r}
# --- Align bigram features with the same train/test rows AND run PCA ---

# 1) Align bigram features with train/test IDs
train_bigram <- train_text %>%
  select(.id, bclass) %>%
  left_join(bigram_features, by = c(".id", "bclass"))

test_bigram <- test_text %>%
  select(.id, bclass) %>%
  left_join(bigram_features, by = c(".id", "bclass"))

# 2) Bigram matrices (drop .id, label)
x_bigram_train <- train_bigram %>%
  select(-.id, -bclass) %>%
  as.matrix()

x_bigram_test <- test_bigram %>%
  select(-.id, -bclass) %>%
  as.matrix()

# 3) Remove zero-variance / constant columns BEFORE PCA
col_sd <- apply(x_bigram_train, 2, sd, na.rm = TRUE)
keep_cols <- which(col_sd > 0)

x_bigram_train2 <- x_bigram_train[, keep_cols, drop = FALSE]
x_bigram_test2  <- x_bigram_test[,  keep_cols, drop = FALSE]

# 4) PCA on bigrams (reduced matrix)
set.seed(123)
bigram_pca <- prcomp(x_bigram_train2, center = TRUE, scale. = TRUE)

k_bigram <- min(50, ncol(bigram_pca$x))  # number of PCs to keep

z_bigram_train <- bigram_pca$x[, 1:k_bigram, drop = FALSE]
z_bigram_test  <- predict(bigram_pca, newdata = x_bigram_test2)[, 1:k_bigram, drop = FALSE]

# 5) Assemble stage-2 train/test data
train_stage2 <- as_tibble(z_bigram_train) %>%
  mutate(
    bclass     = train_text$bclass,
    logit_word = train_logit_word
  )

test_stage2 <- as_tibble(z_bigram_test) %>%
  mutate(
    bclass     = test_text$bclass,
    logit_word = test_logit_word
  )

# 6) Name PCs nicely and put outcome first
pc_names <- paste0("PC_bigram_", 1:k_bigram)
names(train_stage2)[1:k_bigram] <- pc_names
names(test_stage2)[1:k_bigram]  <- pc_names

train_stage2 <- train_stage2 %>% relocate(bclass)
test_stage2  <- test_stage2  %>% relocate(bclass)

```


6. Second-stage logistic regression: logit + bigram PCs

```{r}
# --- Stage-2 logistic regression: logit_word + bigram PCs ---

train_stage2$bclass <- factor(train_stage2$bclass)
test_stage2$bclass  <- factor(test_stage2$bclass,
                              levels = levels(train_stage2$bclass))

stage2_fit <- glm(
  bclass ~ .,
  data   = train_stage2,
  family = binomial()
)

# predict on test set
test_prob_stage2 <- predict(stage2_fit, newdata = test_stage2, type = "response")

b_levels <- levels(train_stage2$bclass)

test_pred_class_stage2 <- factor(
  ifelse(test_prob_stage2 > 0.5, b_levels[2], b_levels[1]),
  levels = b_levels
)

results_stage2 <- tibble(
  bclass      = test_stage2$bclass,
  .pred_class = test_pred_class_stage2
)

acc_stage2 <- accuracy(results_stage2, truth = bclass, estimate = .pred_class)
acc_stage2

```
