---
title: "preliminaryTasks"
output: html_document
---

# Preliminary Task 1: Augmenting HTML to include header
```{r, message = F}
require(tidyverse)
require(tidytext)
require(textstem)
require(rvest)
require(qdapRegex)
require(stopwords)
require(tokenizers)
library(caret)
library(tidymodels)
```

## Updating Pre-processing Functions

Update `parse_fn` to include headers

```{r}
# function to parse html and clean text
parse_fn <- function(.html) {
  read_html(.html) %>%
    html_elements("p, h1, h2, h3, h4, h5, h6") %>%
    html_text2() %>%
    str_c(collapse = " ") %>%
    rm_url() %>%
    rm_email() %>%
    str_remove_all("'") %>%
    str_replace_all(paste(
      c(
        "\n",
        "[[:punct:]]",
        "nbsp",
        "[[:digit:]]",
        "[[:symbol:]]"
      ),
      collapse = "|"
    ), " ") %>%
    str_replace_all("([a-z])([A-Z])", "\\1 \\2") %>%
    tolower() %>%
    str_replace_all("\\s+", " ")
}

# function to apply to claims data
parse_data <- function(.df) {
  out <- .df %>%
    filter(str_detect(text_tmp, "<!")) %>%
    rowwise() %>%
    mutate(text_clean = parse_fn(text_tmp)) %>%
    unnest(text_clean)
  return(out)
}

nlp_fn <- function(parse_data.out) {
  out <- parse_data.out %>%
    unnest_tokens(
      output = token,
      input = text_clean,
      token = "words",
      stopwords = str_remove_all(
        stop_words$word,
        "[[:punct:]]"
      )
    ) %>%
    mutate(token.lem = lemmatize_words(token)) %>%
    filter(str_length(token.lem) > 2) %>%
    count(.id, bclass, token.lem, name = "n") %>%
    bind_tf_idf(
      term = token.lem,
      document = .id,
      n = n
    ) %>%
    pivot_wider(
      id_cols = c(".id", "bclass"),
      names_from = "token.lem",
      values_from = "tf_idf",
      values_fill = 0
    )
  return(out)
}

# clean_df <- claims_test %>%
#   parse_data() %>%
#   select(.id, bclass, text_clean)
load("../data/claims-raw.RData")
claims_clean <- claims_raw %>%
  parse_data()
```

## Logistic PCR

```{r}
text_features <- nlp_fn(claims_clean)
model_df <- text_features %>%
  select(-.id)

model_df$bclass <- as.factor(model_df$bclass)

set.seed(123)
split <- initial_split(model_df, strata = bclass)

train_df <- training(split)
test_df <- testing(split)

pcr_recipe <- recipe(bclass ~ ., data = train_df) %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_predictors()) %>%
  step_pca(all_predictors(), num_comp = 50)

logistic_spec <- logistic_reg(
  mode = "classification",
  penalty = 0.01,
  mixture = 0
) %>%
  set_engine("glmnet")

pcr_wf <- workflow() %>%
  add_recipe(pcr_recipe) %>%
  add_model(logistic_spec)

final_fit <- fit(pcr_wf, train_df)

# Predictions (class)
pred_class <- predict(final_fit, test_df)

# Predictions (probabilities)
pred_prob <- predict(final_fit, test_df, type = "prob")

# Combine with test data
bclass <- as_tibble(test_df$bclass)
results <- bind_cols(bclass, pred_class, pred_prob) %>%
  rename(bclass = value)

# Accuracy
accuracy(results, truth = bclass, estimate = .pred_class)
# 0.616
```


```{r}
text_features <- nlp_fn(claims_clean)
model_df <- text_features %>%
  select(-.id)
```

```{r}
model_df$bclass <- as.factor(model_df$bclass)

set.seed(123)
split <- initial_split(model_df, strata = bclass)
train_df <- training(split)
test_df <- testing(split)

# Recipe with fixed PCA components (no tuning)
pcr_recipe <- recipe(bclass ~ ., data = train_df) %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_predictors()) %>%
  step_pca(all_predictors(), num_comp = 50)

# Logistic regression model
logistic_spec <- logistic_reg(
  mode    = "classification",
  penalty = 0.01,
  mixture = 0
) %>%
  set_engine("glmnet")

# Workflow
pcr_wf <- workflow() %>%
  add_recipe(pcr_recipe) %>%
  add_model(logistic_spec)

# Fit the model directly
final_fit <- fit(pcr_wf, train_df)

# Predictions
pred_class <- predict(final_fit, test_df)
pred_prob <- predict(final_fit, test_df, type = "prob")

# Combine with test labels
results <- bind_cols(
  bclass = test_df$bclass,
  pred_class,
  pred_prob
) %>%
  rename(bclass = bclass)

# Accuracy
accuracy(results, truth = bclass, estimate = .pred_class)
```

## Preliminary task 2 

1. Bigram tokenization and TF–IDF

Add a bigram nlp function that parallels your nlp_fn():

```{r}
# --- Task 2: Bigram tokenization and TF–IDF ---

nlp_fn_bigrams <- function(parse_data.out) {
  parse_data.out %>%
    unnest_tokens(
      output = bigram,
      input  = text_clean,
      token  = "ngrams",
      n      = 2
    ) %>%
    # optional cleanup: drop bigrams with digits or punctuation
    filter(
      !str_detect(bigram, "[0-9]"),
      !str_detect(bigram, "[[:punct:]]")
    ) %>%
    count(.id, bclass, bigram, name = "n") %>%
    bind_tf_idf(
      term     = bigram,
      document = .id,
      n        = n
    ) %>%
    pivot_wider(
      id_cols     = c(".id", "bclass"),
      names_from  = "bigram",
      values_from = "tf_idf",
      values_fill = 0
    )
}

# recompute word-level features to be explicit (safe even if already done)
text_features <- nlp_fn(claims_clean)

# bigram-level features
bigram_features <- nlp_fn_bigrams(claims_clean)
```

2. Create a consistent train/test split by ID
```{r}
# IDs + label for splitting
set.seed(123)
split2 <- initial_split(text_features, strata = bclass)

train_text <- training(split2)
test_text <- testing(split2)

# data frames for word-level modeling (drop .id)
train_word_df <- train_text %>% select(-.id)
test_word_df <- test_text %>% select(-.id)

train_word_df$bclass <- as.factor(train_word_df$bclass)
test_word_df$bclass <- as.factor(test_word_df$bclass)
```
3. Word-only logistic PCR (baseline for Task 2)

```{r}
# --- Word-only logistic PCR (baseline) ---

# --- Word-only logistic PCR (baseline with tuning) ---

set.seed(123)
pcr_recipe_word <- recipe(bclass ~ ., data = train_word_df) %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_predictors()) %>%
  # make num_comp tunable instead of fixed at 50
  step_pca(all_predictors(), num_comp = tune())

logistic_spec <- logistic_reg(
  mode    = "classification",
  penalty = 0.01,
  mixture = 0
) %>%
  set_engine("glmnet")

pcr_wf_word <- workflow() %>%
  add_recipe(pcr_recipe_word) %>%
  add_model(logistic_spec)

set.seed(123)
folds_word <- vfold_cv(train_word_df, v = 5, strata = bclass)

# grid of PCA components to try
pca_grid <- tibble(num_comp = c(10, 20, 50, 100, 200))

tune_res_word <- tune_grid(
  pcr_wf_word,
  resamples = folds_word,
  grid      = pca_grid,
  metrics   = metric_set(accuracy)
)

best_pcs_word <- select_best(tune_res_word, metric = "accuracy")
best_pcs_word

final_wf_word <- finalize_workflow(pcr_wf_word, best_pcs_word)

final_fit_word <- fit(final_wf_word, train_word_df)

# predictions on test set (baseline)
pred_class_word <- predict(final_fit_word, test_word_df)

results_word <- bind_cols(
  bclass = test_word_df$bclass,
  pred_class_word
) %>%
  rename(bclass = bclass)

acc_word <- accuracy(results_word, truth = bclass, estimate = .pred_class)
acc_word
```

4. Get predicted log-odds (logit) from the word model

```{r}
# --- Predicted log-odds from word-only model ---

# training probabilities
train_prob_word <- predict(final_fit_word, train_word_df, type = "prob")$.pred_1
test_prob_word <- predict(final_fit_word, test_word_df, type = "prob")$.pred_1

# safe logit function (avoid log(0))
logit <- function(p) {
  eps <- 1e-6
  p <- pmin(pmax(p, eps), 1 - eps)
  log(p / (1 - p))
}

train_logit_word <- logit(train_prob_word)
test_logit_word <- logit(test_prob_word)
```

5. PCA on bigram TF–IDF and build stage-2 design matrices

```{r}
# --- Align bigram features with the same train/test rows ---

train_bigram <- train_text %>%
  select(.id, bclass) %>%
  left_join(bigram_features, by = c(".id", "bclass"))

test_bigram <- test_text %>%
  select(.id, bclass) %>%
  left_join(bigram_features, by = c(".id", "bclass"))

# bigram matrices (drop .id, label)
x_bigram_train <- train_bigram %>%
  select(-.id, -bclass) %>%
  as.matrix()

x_bigram_test <- test_bigram %>%
  select(-.id, -bclass) %>%
  as.matrix()

# PCA on bigrams
set.seed(123)
bigram_pca <- prcomp(x_bigram_train, center = TRUE, scale. = TRUE)

k_bigram <- min(50, ncol(bigram_pca$x)) # number of PCs to keep

z_bigram_train <- bigram_pca$x[, 1:k_bigram, drop = FALSE]
z_bigram_test <- predict(bigram_pca, newdata = x_bigram_test)[, 1:k_bigram, drop = FALSE]

# assemble stage-2 train/test data
train_stage2 <- as_tibble(z_bigram_train) %>%
  mutate(
    bclass     = train_text$bclass,
    logit_word = train_logit_word
  )

test_stage2 <- as_tibble(z_bigram_test) %>%
  mutate(
    bclass     = test_text$bclass,
    logit_word = test_logit_word
  )

# name PCs nicely and put outcome first
pc_names <- paste0("PC_bigram_", 1:k_bigram)
names(train_stage2)[1:k_bigram] <- pc_names
names(test_stage2)[1:k_bigram] <- pc_names

train_stage2 <- train_stage2 %>% relocate(bclass)
test_stage2 <- test_stage2 %>% relocate(bclass)
```

6. Second-stage logistic regression: logit + bigram PCs

```{r}
# --- Stage-2 logistic regression: logit_word + bigram PCs ---

train_stage2$bclass <- factor(train_stage2$bclass)
test_stage2$bclass <- factor(test_stage2$bclass,
  levels = levels(train_stage2$bclass)
)

stage2_fit <- glm(
  bclass ~ .,
  data   = train_stage2,
  family = binomial()
)

# predict on test set
test_prob_stage2 <- predict(stage2_fit, newdata = test_stage2, type = "response")

b_levels <- levels(train_stage2$bclass)

test_pred_class_stage2 <- factor(
  ifelse(test_prob_stage2 > 0.5, b_levels[2], b_levels[1]),
  levels = b_levels
)

results_stage2 <- tibble(
  bclass      = test_stage2$bclass,
  .pred_class = test_pred_class_stage2
)

acc_stage2 <- accuracy(results_stage2, truth = bclass, estimate = .pred_class)
acc_stage2
```


```{r}
getwd()
```

# Preliminary Task 3: Neural Network Model

```{r, message = FALSE}
library(keras)
library(tensorflow)
```

## Data partitioning for neural network

We'll use the `claims_clean` data from Task 1 (already includes headers).

```{r}
set.seed(110122)
partitions <- claims_clean %>%
  initial_split(prop = 0.8, strata = bclass)

# Extract training data
train_text <- training(partitions) %>%
  pull(text_clean)
train_labels <- training(partitions) %>%
  pull(bclass) %>%
  as.numeric() - 1

# Extract test data
test_text <- testing(partitions) %>%
  pull(text_clean)
test_labels <- testing(partitions) %>%
  pull(bclass) %>%
  as.numeric() - 1

cat("Training set size:", length(train_labels), "\n")
cat("Test set size:", length(test_labels), "\n")
cat("Training set claim proportion:", mean(train_labels), "\n")
```

## Text vectorization layer

Create a preprocessing layer using TF-IDF with a vocabulary of 5,000 tokens.

```{r}
# Create a preprocessing layer with TF-IDF
preprocess_layer <- layer_text_vectorization(
  standardize = NULL,
  split = "whitespace",
  ngrams = NULL,
  max_tokens = 5000,
  output_mode = "tf_idf"
)

# Adapt the preprocessing layer to the training data
preprocess_layer %>% adapt(train_text)

cat("Vocabulary size:", preprocess_layer$vocabulary_size(), "\n")
```

## Define neural network architecture

Build a feedforward neural network with three hidden layers and dropout regularization.

```{r}
model <- keras_model_sequential() %>%
  preprocess_layer() %>%
  layer_dropout(0.3) %>%
  layer_dense(units = 64, activation = "relu") %>%
  layer_dropout(0.3) %>%
  layer_dense(units = 32, activation = "relu") %>%
  layer_dropout(0.2) %>%
  layer_dense(units = 16, activation = "relu") %>%
  layer_dropout(0.2) %>%
  layer_dense(1) %>%
  layer_activation(activation = "sigmoid")

summary(model)
```

## Configure and train the model

Use Adam optimizer with binary cross-entropy loss, and employ early stopping to prevent overfitting.

```{r}
# Compile model
model %>% compile(
  loss = "binary_crossentropy",
  optimizer = optimizer_adam(learning_rate = 0.001),
  metrics = c("binary_accuracy")
)

# Train with early stopping
history <- model %>%
  fit(
    train_text,
    train_labels,
    validation_split = 0.2,
    epochs = 20,
    batch_size = 32,
    verbose = 2,
    callbacks = list(
      callback_early_stopping(
        monitor = "val_loss",
        patience = 5,
        restore_best_weights = TRUE
      )
    )
  )
```

## Visualize training history

```{r}
plot(history)
```

## Evaluate on test set

```{r}
# Get predictions
test_pred_prob <- model %>% predict(test_text)
test_pred_class <- ifelse(test_pred_prob > 0.5, 1, 0)

# Calculate accuracy
test_accuracy <- mean(test_pred_class == test_labels)

cat("\n============================================================\n")
cat("TEST SET RESULTS\n")
cat("============================================================\n")
cat("Test Accuracy:", round(test_accuracy, 4), "\n")
cat("Test Accuracy (%):", round(test_accuracy * 100, 2), "%\n\n")

# Confusion matrix
confusion_mat <- table(Predicted = test_pred_class, Actual = test_labels)
cat("Confusion Matrix:\n")
print(confusion_mat)
cat("\n")

# Additional metrics
tp <- confusion_mat[2, 2]
fp <- confusion_mat[2, 1]
tn <- confusion_mat[1, 1]
fn <- confusion_mat[1, 2]

precision <- tp / (tp + fp)
recall <- tp / (tp + fn)
f1_score <- 2 * (precision * recall) / (precision + recall)

cat("Precision:", round(precision, 4), "\n")
cat("Recall:", round(recall, 4), "\n")
cat("F1-Score:", round(f1_score, 4), "\n\n")

# Compare with Task 1 baseline
baseline_accuracy <- 0.616
cat("Comparison with Task 1 (Logistic PCR):\n")
cat("Baseline accuracy:", baseline_accuracy, "\n")
cat("NN accuracy:", round(test_accuracy, 4), "\n")
if (test_accuracy >= baseline_accuracy) {
  cat("✓ Neural network MEETS or EXCEEDS baseline!\n")
  cat(
    "Improvement:", round((test_accuracy - baseline_accuracy) * 100, 2),
    "percentage points\n"
  )
} else {
  cat("✗ Neural network below baseline\n")
}
```

## Save model and predictions

```{r}
# Save the trained model
model_path <- "../results/task3-nn-model"
save_model_tf(model, model_path)
cat("Model saved to:", model_path, "\n")

# Save predictions
test_ids <- testing(partitions) %>% pull(.id)
predictions_df <- tibble(
  .id = test_ids,
  bclass_actual = test_labels,
  bclass_pred_prob = as.vector(test_pred_prob),
  bclass_pred_class = as.vector(test_pred_class)
)

save(predictions_df, file = "../results/task3-predictions.RData")
cat("Predictions saved to: ../results/task3-predictions.RData\n")
```