---
title: "Random Forest Model"
author: "Kaeya Mehta"
date: "2025-11-19"
output: html_document
---

# Libraries 

```{r}
require(tidyverse)
require(tidytext)
require(textstem)
require(rvest)
require(qdapRegex)
require(stopwords)
require(tokenizers)
library(caret)
library(tidymodels)
library(dplyr)
library(randomForest)
```


# Preprocessing

Load raw data
```{r}
load("/Users/kaeya/PSTAT 197A/module-2-claims-data-table-12-1/data/claims-raw.RData")
load("/Users/kaeya/PSTAT 197A/module-2-claims-data-table-12-1/data/claims-test.RData")
```

```{r}
view(claims_test)
```


Functions to preprocess
```{r}
# function to parse html and clean text
parse_fn <- function(.html) {
  read_html(.html) %>%
    html_elements("p, h1, h2, h3, h4, h5, h6") %>%
    html_text2() %>%
    str_c(collapse = " ") %>%
    rm_url() %>%
    rm_email() %>%
    str_remove_all("'") %>%
    str_replace_all(paste(
      c(
        "\n",
        "[[:punct:]]",
        "nbsp",
        "[[:digit:]]",
        "[[:symbol:]]"
      ),
      collapse = "|"
    ), " ") %>%
    str_replace_all("([a-z])([A-Z])", "\\1 \\2") %>%
    tolower() %>%
    str_replace_all("\\s+", " ")
}

# function to apply to claims data
parse_data <- function(.df) {
  out <- .df %>%
    filter(str_detect(text_tmp, "<!")) %>%
    rowwise() %>%
    mutate(text_clean = parse_fn(text_tmp)) %>%
    unnest(text_clean)
  return(out)
}

```


Cleaned data
```{r}
set.seed(123)
claims_split <- claims_raw %>%
  initial_split(prop = 0.8)
claims_train_cv <- training(claims_split)
claims_test_cv <- testing(claims_split)

claims_clean_train <- parse_data(claims_train_cv)
claims_clean_test <- parse_data(claims_test_cv)
```


Tokenization for Train - Binary
```{r}
stpwrd <- stop_words %>%
  pull(word) %>%
  str_remove_all("[[:punct:]]")

tokens_all <- claims_clean_train %>%
  select(.id, bclass, mclass, text_clean) %>%    
  unnest_tokens(
    output = token,
    input  = text_clean,
    token  = "words",
    stopwords = stpwrd
  ) %>%
  mutate(token = lemmatize_words(token)) %>%     
  filter(str_length(token) > 2)  

tf_counts <- tokens_all %>%
  dplyr::count(.id, bclass, token, name = "n")

claims_tfidf_binary <- tf_counts %>%
  bind_tf_idf(
    term     = token,
    document = .id,
    n        = n
  )

claims_training_binary <- claims_tfidf_binary %>%
  tidyr::pivot_wider(
    id_cols     = c(.id, bclass),
    names_from  = token,
    values_from = tf_idf,
    values_fill = 0
  )

colnames(claims_training_binary) <- make.names(colnames(claims_training_binary))

token_cols <- setdiff(names(claims_training_binary), c(".id", "bclass"))

nonzero_counts <- colSums(claims_training_binary[, token_cols] != 0)
keep_tokens <- names(nonzero_counts)[nonzero_counts >= 5]

vocabbin <- keep_tokens

claims_training_small <- claims_training_binary %>%
  select(bclass, all_of(keep_tokens))

claims_training_small
```

Tokenization for Test - Binary
```{r}
stpwrd <- stop_words %>%
  pull(word) %>%
  str_remove_all("[[:punct:]]")

tokens_all <- claims_clean_test %>%
  select(.id, bclass, mclass, text_clean) %>%    
  unnest_tokens(
    output = token,
    input  = text_clean,
    token  = "words",
    stopwords = stpwrd
  ) %>%
  mutate(token = lemmatize_words(token)) %>%     
  filter(str_length(token) > 2,
         token %in% vocabbin)  

tf_counts <- tokens_all %>%
  dplyr::count(.id, bclass, token, name = "n")

claims_tfidf_binary <- tf_counts %>%
  bind_tf_idf(
    term     = token,
    document = .id,
    n        = n
  )

claims_testing_binary <- claims_tfidf_binary %>%
  tidyr::pivot_wider(
    id_cols     = c(.id, bclass),
    names_from  = token,
    values_from = tf_idf,
    values_fill = 0
  )

missing_in_test <- setdiff(vocabbin, names(claims_testing_binary))
for (tok in missing_in_test) {
  claims_testing_binary[[tok]] <- 0
}

claims_testing_small <- claims_testing_binary %>%
  select(bclass, all_of(vocabbin))

claims_testing_small

```

Run model - Binary
```{r}
set.seed(123)

colnames(claims_training_small) <- make.names(colnames(claims_training_small))

claims_training_small$bclass <- factor(claims_training_small$bclass)

rf.claims_training_binary = randomForest(bclass ~ ., data=claims_training_small, importance=TRUE, ntree = 100)
```

Predict model - Binary
```{r}
colnames(claims_testing_small) <- make.names(colnames(claims_testing_small))
preds <- predict(rf.claims_training_binary, newdata = claims_testing_small)
```

Accuracy - Binary
```{r}
conftable <- table(preds, claims_testing_small$bclass)
conftable

accuracy <- sum(diag(conftable)) / sum(conftable)
accuracy
```




Tokenization for Train - Multiple Classification

```{r}
stpwrd <- stop_words %>%
  pull(word) %>%
  str_remove_all("[[:punct:]]")

tokens_all <- claims_clean_train %>%
  select(.id, bclass, mclass, text_clean) %>%    
  unnest_tokens(
    output = token,
    input  = text_clean,
    token  = "words",
    stopwords = stpwrd
  ) %>%
  mutate(token = lemmatize_words(token)) %>%     
  filter(str_length(token) > 2)  

tf_counts <- tokens_all %>%
  dplyr::count(.id, mclass, token, name = "n")

claims_tfidf_multiple <- tf_counts %>%
  bind_tf_idf(
    term     = token,
    document = .id,
    n        = n
  )

claims_training_multiple <- claims_tfidf_multiple %>%
  tidyr::pivot_wider(
    id_cols     = c(.id, mclass),
    names_from  = token,
    values_from = tf_idf,
    values_fill = 0
  )

token_cols <- setdiff(names(claims_training_multiple), c(".id", "mclass"))

nonzero_counts <- colSums(claims_training_multiple[, token_cols] != 0)
keep_tokens <- names(nonzero_counts)[nonzero_counts >= 5]

vocab1 <- keep_tokens

claims_training_small <- claims_training_multiple %>%
  select(mclass, all_of(keep_tokens))

claims_training_small
```


Tokenization for Test - Multiple
```{r}
stpwrd <- stop_words %>%
  pull(word) %>%
  str_remove_all("[[:punct:]]")

tokens_all <- claims_clean_test %>%
  select(.id, bclass, mclass, text_clean) %>%    
  unnest_tokens(
    output = token,
    input  = text_clean,
    token  = "words",
    stopwords = stpwrd
  ) %>%
  mutate(token = lemmatize_words(token)) %>%     
  filter(str_length(token) > 2,
         token %in% vocab)  

tf_counts <- tokens_all %>%
  dplyr::count(.id, mclass, token, name = "n")

claims_tfidf_multiple <- tf_counts %>%
  bind_tf_idf(
    term     = token,
    document = .id,
    n        = n
  )

claims_testing_multiple <- claims_tfidf_multiple %>%
  tidyr::pivot_wider(
    id_cols     = c(.id, mclass),
    names_from  = token,
    values_from = tf_idf,
    values_fill = 0
  )

missing_in_test <- setdiff(vocab, names(claims_testing_multiple))
for (tok in missing_in_test) {
  claims_testing_multiple[[tok]] <- 0
}

claims_testing_small <- claims_testing_multiple %>%
  select(mclass, all_of(vocab))

claims_testing_small
```


Run model - Multiple
```{r}
set.seed(123)

colnames(claims_training_small) <- make.names(colnames(claims_training_small))

claims_training_small$mclass <- factor(claims_training_small$mclass)

rf.claims_training_multiple = randomForest(mclass ~ ., data=claims_training_small, importance=TRUE, ntree = 100)
```

Predict model - Multiple
```{r}
colnames(claims_testing_small) <- make.names(colnames(claims_testing_small))
preds <- predict(rf.claims_training_multiple, newdata = claims_testing_small)
```

Accuracy - Multiple
```{r}
conftable <- table(preds, claims_testing_small$mclass)
conftable

accuracy <- sum(diag(conftable)) / sum(conftable)
accuracy
```



Preprocessing for test data
```{r}
claims_clean_testdf <- parse_data(claims_test)

stpwrd <- stop_words %>%
  pull(word) %>%
  str_remove_all("[[:punct:]]")

tokens_all <- claims_clean_testdf %>%
  select(.id, text_clean) %>%    
  unnest_tokens(
    output = token,
    input  = text_clean,
    token  = "words",
    stopwords = stpwrd
  ) %>%
  mutate(token = lemmatize_words(token)) %>%     
  filter(str_length(token) > 2,
         token %in% vocabbin)  

tf_counts <- tokens_all %>%
  dplyr::count(.id, token, name = "n")

claims_tfidf_binary <- tf_counts %>%
  bind_tf_idf(
    term     = token,
    document = .id,
    n        = n
  )

claims_testing_binary <- claims_tfidf_binary %>%
  tidyr::pivot_wider(
    id_cols     = c(.id),
    names_from  = token,
    values_from = tf_idf,
    values_fill = 0
  )

missing_in_test <- setdiff(vocabbin, names(claims_testing_binary))
for (tok in missing_in_test) {
  claims_testing_binary[[tok]] <- 0
}

claims_testing_small <- claims_testing_binary %>%
  select(all_of(vocabbin))

claims_testing_small
```

```{r}
predsbin <- predict(rf.claims_training_binary, newdata = claims_testing_small)
predsmult <- predict(rf.claims_training_multiple, newdata = claims_testing_small)

pred_df <- tibble(
  .id = claims_testing_binary$.id,
  bclass.pred = predsbin,
  mclass.pred = predsmult
)

pred_df
```

```{r}
# Save predictions
save(pred_df, file = "/Users/kaeya/PSTAT 197A/module-2-claims-data-table-12-1/results/preds-rf.RData")

# Save models
saveRDS(rf.claims_training_binary, "/Users/kaeya/PSTAT 197A/module-2-claims-data-table-12-1/results/rf-binary.rds")
saveRDS(rf.claims_training_multiple, "/Users/kaeya/PSTAT 197A/module-2-claims-data-table-12-1/results/rf-multiclass.rds")
```





